---
title: "Gapminder-dlpyr"
author: "Bea"
date: "19/06/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

##TIDYING DATA WITH dplyr

Loading tidyverse
```{r}
library(tidyverse)
```

Load data
```{r data}
gapminder <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")

head(gapminder)
tail(gapminder)
str(gapminder) #structure of the data, best if it is a data.frame
names(gapminder)
ncol(gapminder)
nrow(gapminder)
```

Start using dplyr functions
```{r functions}
filter(gapminder, lifeExp < 35)
filter(gapminder, country == "Mexico", year == 2002)

```

Filter function again
```{r filter}
life_exp <- filter(gapminder, country == "Brazil", year > 1986)
#life_exp <- filter(gapminder, country == "Brazil", year > 1986 | year < 2008)
mean(life_exp$lifeExp)
```

Select function
```{r select}
select(gapminder, year, country, lifeExp)
#select(gapminder, -continent, -gdpPercap) #The same but using -
```

Using filter and select together
```{r mixed}
gap_cambodia <- filter(gapminder, country == "Cambodia")
gap_cambodia2 <- select(gap_cambodia, -continent, -lifeExp)
```

Using the pipe operator to make things easier: This is going to save my life

```{r way1}
gapminder %>% head()
gap_cambodia <- filter(gapminder, country == "Cambodia") %>% 
select(-continent, -lifeExp)

```
```{r way2}
gap_cambodia <- gapminder %>% filter(country == "Cambodia") %>% 
select(-continent, -lifeExp) 
```

Learn `mutate()` which adds a new column

```{r}
gap2 <- gapminder %>% 
  mutate(gdp = pop  * gdpPercap)
```

Exercise
```{r asian_pop}
gapminder %>% filter(continent == "Asia", year == 2007) %>%
  mutate(pop_thousands = pop/1000) %>% 
  select(country, year, pop_thousands)
 
```

Using the `group_by()` funtion
```{r}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  mutate(cont_pop = sum(pop))

```

Using summarize to reduce the number of data
```{r}
gapminder %>% 
  filter(year == 2002) %>% 
  group_by(continent) %>% 
  summarize(cont_pop = sum(pop))

```

Ungroup to undo the grouping, and we always need to ungroup at the end, because R still thinks that your data frame should be grouped
```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  ungroup()
```

Using 'arrange()` to arrange the columns in the data frame. If a "-" is used, it means to arranged thw other way

```{r}
gapminder %>% 
  group_by(continent, year) %>% 
  summarize(cont_pop = sum(pop)) %>% 
  arrange(-year)
```

New exercise for arrange

```{r}

gapminder %>%
  mutate(gdp = pop*gdpPercap) %>% 
  group_by(continent) %>% 
  mutate(max_gdp = max(gdp)) %>% #calculate the maximum gdp using the max() function
  filter(gdp == max_gdp) #we only take the maximum for each country

```

