---
title: "Graphics with ggplot2"
author: "Bea"
date: "11/21/2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Learning ggplot2

Load Tidyverse
```{r tidyverse}
library(tidyverse) ## do this every time you restart R and need it 
```

We're learning ggplot2 It's going to be amazing. 

```{r data}

ohi_data <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_global_data.csv")

```
We start!!

```{r start ggplot} 
ggplot(ohi_data, aes(x=georegion_one, y=OHI_score)) + 
  geom_point()

ggplot(ohi_data, aes(x=georegion_one, y=OHI_score)) + 
  geom_jitter(width=0.2)

ggplot(ohi_data, aes(x=georegion_one, y=OHI_score)) + 
  geom_count()
```

Boxplot
```{r boxplot}
ggplot(ohi_data, aes(x=georegion_one)) + 
  geom_bar()

```

Histogram
```{r histogram}
ggplot(ohi_data, aes(x=HDI)) + 
  geom_histogram()

```

Multiple geoms exploration

```{r multiple geoms}

ohi_summary <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/OHI_scores_georegion_summary.csv")

ggplot(ohi_summary, aes(x=georegions, y=OHI_score_average)) +
  geom_bar(stat = "identity") +
  geom_jitter(data= ohi_data, aes(x=georegion_one, y= OHI_score))

```

Are these two approaches the same?

```{r two same plots}

ggplot(data = ohi_data, aes(y=OHI_score, x = HDI, color=georegion_one)) +
  geom_point()
  
ggplot(data = ohi_data) +
  geom_point(aes(y = OHI_score, x = HDI, color=georegion_one))

```
These two approaches result in the same plot here, but there could be downstream effects as more layers are added.

The arguments within aes() link variables in the dataframe to some aspect of plot appearance. As we have discussed, x and y describe the axes, but other arguments can be added to describe a z variable (e.g. size or color or shape of points). Here are some examples of aes arguments:

color - color of lines/points
fill - color within polygons
label - name
linetype - type of line
shape - style of point
alpha - transparency (0-1)
size - size of shape

```{r features test}

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, size = coastal_pop)) + 
  geom_point()

```
In that code we add a third variable by modifiying the size of the points in the data by coastal_pop variable

```{r features test2}

ggplot(data = ohi_data, aes(x = OHI_score, y = HDI, size = coastal_pop, color =georegion_one)) + geom_point()

```
Make our plots way prettier. For that we use themes.

```{r themes}

source('https://raw.githubusercontent.com/OHI-Science/ohiprep/master/src/R/scatterTheme.txt')

ggplot(ohi_data, aes(x=OHI_score, y= HDI)) +
  geom_point() +
  scatterTheme

ggplot(ohi_data, aes(x=OHI_score, y= HDI, color=georegion_one)) +
  geom_point() +
  labs(y= "OHI score, 2017", x="Human development index", title= "Human development is good for oceans!", color = "Georegions")

```
Color stuff!!!!! Using RColorBrewer to personalize the color palette

```{r brewer}
library("RColorBrewer")
library("colorspace")

display.brewer.all()
my_palette <- brewer.pal(n=9, "YlOrRd")

```
There are different functions to color the plots depending on the type of variable you have (discrete or continuous). 

```{r variables}

ggplot(ohi_data, aes(x=OHI_score, y=OHI_trend, color=HDI)) +
  geom_point( size = 3) +
  scale_color_gradientn(colors=my_palette)

#Assign the color using a vector with discrete colors

ggplot(ohi_data, aes(x=OHI_score, y=OHI_trend, color=HDI)) +
  geom_point( size = 3) +
  scale_color_gradientn(colors=c("blue", "red"))

my_palette<-brewer.pal(n=12, "Set3")
ggplot(ohi_data, aes(x=OHI_score, y=OHI_trend, color=georegion_one)) +
  geom_point( size = 3) +
  scale_color_manual(values = my_palette)

```

##TIDYING DATA WITH dplyr


